### 2016-03-21 manuscript.Rmd
Add ACM SIG proceedings citation style file

---
title: "Rethinking how students learn to design programs"
author: "Blinded for Review"
date: "March 20, 2016"
output:
  pdf_document:
    keep_tex: true
    number_sections: true
    highlight: "tango"
    toc: true
    md_extensions: +implicit_figures +grid_tables +fenced_code_blocks +fenced_code_attributes
    citation_package: 'none'
    pandoc_args: [
      --bibliography=bibliography.bib,
      --csl=acm-sig-proceedings-long-author-list.csl,
      --filter=pandoc-citeproc
    ]
  word_document:
    highlight: "tango"
    toc: true
    md_extensions: +implicit_figures +grid_tables +fenced_code_blocks +fenced_code_attributes
    pandoc_args: [
      --bibliography=bibliography.bibtex,
      --csl=acm-sig-proceedings-long-author-list.csl,
      --filter=pandoc-citeproc
    ]---