### 2016-03-24 manuscript.Rmd
Add sample files for ACM style

---
title: "Rethinking how students learn to design programs"
author: "Blinded for Review"
date: "March 20, 2016"
abstract: This is the abstract
documentclass: sig-alternate
geometry: paperwidth=8.5in paperheight=11in
includes:
output:
  pdf_document:
    md_extensions: +implicit_figures +grid_tables +fenced_code_blocks +fenced_code_attributes +raw_tex
    keep_tex: true
  word_document:
    reference_docx: pubform.docx
    highlight: "tango"
    md_extensions: +implicit_figures +grid_tables +fenced_code_blocks +fenced_code_attributes
    pandoc_args: [
      --smart,
      --bibliography=bibliography.bib,
      --csl=acm-sig-proceedings,
      --filter=pandoc-citeproc
    ]
---

# Section 1