### 2016-04-01 manuscript.Rmd
Convert `geometry` to proper YAML
- Remove `includes` YAML field
- Remove word-document RMarkdown output

---
title: "Rethinking how students learn to design programs"
author: "Author Blinded for Review"
date: "March 20, 2016"
abstract: This is the complete abstract
documentclass: sig-alternate
geometry:
  - paperwidth=8.5in
  - paperheight=11in
template: sig-alternate-modified-for-pandoc.tex
output:
  pdf_document:
    md_extensions: +implicit_figures +grid_tables +fenced_code_blocks +fenced_code_attributes +raw_tex
    keep_tex: true
    highlight: "tango"
    pandoc_args: [
      --smart,
      --template=sig-alternate-modified-for-pandoc
    ]
---

# Section 1

## Subsection