### 2016-04-22 RebeccasCode/RebeccasOneStopFlight.Rmd

---
title: "Rebecca’s One-Stop Flight Code"
output:
  pdf_document:
    highlight: kate
geometry: paperwidth=40in, paperheight=72in
---

```c
//"Rebecca Wells"
//1 stop flight function

#include <stdio.h>
#include <string.h>
//#include "airports.h"

int one_stop_flight(int choice)
{
  int route_id, id, input, flight_number, d_hour, d_mins, a_hour, a_mins, d_hour_stop, d_mins_stop, a_hour_stop, a_mins_stop, z, layover, b;
  char departure_input[4], arrival_input[4], stop_code[4], days[7], departure_city[50], arrival_city[50], stop_city[50], arrival_time[7], departure_time[7], a_letter, d_letter, travel_days[10], airport_code[4], city[50], arrival_airport[4], departure_airport[4], a_letter_stop, d_letter_stop;
  FILE *airports, *flights, *routes;

  printf("Enter departure and arrival airport codes: ");
  scanf("%s %s", departure_input, arrival_input);
  flights = fopen("flights.txt", "r");
  airports = fopen("airports.txt", "r");
  routes = fopen("routes.txt", "r");

  while(fscanf(airports, "%s  %50[^\n]", airport_code, city) != EOF)   //scan for the departure code but not to the arrival code
    {
      if(airport_code != departure_input || airport_code != arrival_input)
        {
          while(fscanf(routes, "%d.  %s %s", &route_id, departure_airport, arrival_airport) != EOF)  //finds the name of the city of the departure airport
            {
              if(departure_input[0]==departure_airport[0] && departure_input[1]==departure_airport[1] && departure_input[2]==departure_airport[2])
                {
                  if(arrival_airport[0] != arrival_input[0] || arrival_airport[1] != arrival_input[1] || arrival_airport[2] != arrival_input[2])
                    {
                      strcpy(stop_code, arrival_airport);
                      rewind(routes);
                      departure_airport[0]=' ';
                      departure_airport[1]=' ';
                      departure_airport[2]=' ';
                      arrival_airport[0]=' ';
                      arrival_airport[1]=' ';
                      arrival_airport[2]=' ';
                      while(fscanf(routes, "%d.  %s %s", &route_id, departure_airport, arrival_airport) != EOF)  //now searches for the place where the stop is the
                        {                                                                                      //depart airport and the arrival input is correct
                          if(stop_code[0]==departure_airport[0] && stop_code[1]==departure_airport[1] && stop_code[2]==departure_airport[2] && arrival_input[0]==arrival_airport[0] && arrival_input[1]==arrival_airport[1] && arrival_input[2]==arrival_airport[2])
                            {
                              id=route_id;
                              printf("Enter a day of the week (1-7, or 0 for all days):");
                              scanf("%d", &input);

                              //now depending on the day entered, program is suppose to run through each option that was entered, find where the id is similar, then
                              //check the days, finally check the time that is involved in the stop to make sure the layover is the correct time
                              //end with printing out the message if you find a suitable one stop flight

                              if(input==1)
                                {
                                  while(fscanf(flights, "%d %d %d:%d%c %d:%d%c %s", flight_number, route_id, d_hour, d_mins, d_letter, a_hour, a_mins, a_letter, days) != EOF)
                                    {
                                      if(id==route_id)
                                        {
                                          while(z<7)
                                            {
                                              if(days[z]==1)
                                                {
                                                  //if(check_days() == TRUE)
                                                    {
                                                      // check_time();
                                                      if(layover >= 60 && layover <= 120)
                                                        {
                                                          printf("Flight %d (%s)\n    %d:%d%c.m. %s (%s)\n    %d:%d%c.m. %s (%s)\n", flight_number, travel_days, d_hour, d_mins, d_letter, departure_city, departure_input, a_hour_stop, a_mins_stop, a_letter_stop, stop_city, stop_code);
                                                          printf("Flight %d (%s)\n    %d:%d.m. %s, %s (%s)\n    %d:%d%c.m. %s (%s)\n", flight_number, (days), d_hour_stop, d_mins_stop, d_letter_stop, stop_city, stop_code, a_hour, a_mins, a_letter, arrival_city, arrival_input);
                                                        }
                                                    }
                                                }
                                            }
                                        }
                                    }
                                }

                              //if day entered is 2

                              if(input==2)
                                {
                                  while(fscanf(flights, "%d %d %d:%d%c %d:%d%c %s", flight_number, route_id, d_hour, d_mins, d_letter, a_hour, a_mins, a_letter, days) != EOF)
                                    {
                                      if(id==route_id)
                                        {
                                          while(z<7)
                                            {
                                              if(days[z]==2)
                                                {
                                                 // if(check_days() == TRUE)
                                                    {
                                                      //check_time();
                                                      if(layover >= 60 && layover <= 120)
                                                        {
                                                          printf("Flight %d (%s)\n    %d:%d%c.m. %s (%s)\n    %d:%d%c.m. %s (%s)\n", flight_number, travel_days, d_hour, d_mins, d_letter, departure_city, departure_input, a_hour_stop, a_mins_stop, a_letter_stop, stop_city, stop_code);
                                                          printf("Flight %d (%s)\n    %d:%d%c.m. %s (%s)\n    %d:%d%c.m. %s (%s)\n", flight_number, (days), d_hour_stop, d_mins_stop, d_letter_stop, stop_city, stop_code, a_hour, a_mins, a_letter, arrival_city, arrival_input);
                                                        }
                                                    }  
                                                }
                                            }
                                        }             
                                    }
                                }

                              //if day entered is 3

                              if(input==3)
                                {
                                  while(fscanf(flights, "%d %d %d:%d%c %d:%d%c %s", flight_number, route_id, d_hour, d_mins, d_letter, a_hour, a_mins, a_letter, days) != EOF)
                                    {
                                      if(id==route_id)
                                        {
                                          while(z<7)
                                            {
                                              if(days[z]==3)
                                                {
                                                  //if(check_days() == TRUE)
                                                    {
                                                      //check_time();
                                                      if(layover >= 60 && layover <= 120)
                                                        {
                                                          printf("Flight %d (%s)\n    %d:%d%c.m. %s (%s)\n    %d:%d%c.m. %s (%s)\n", flight_number, travel_days, d_hour, d_mins, d_letter, departure_city, departure_input, a_hour_stop, a_mins_stop, a_letter_stop, stop_city, stop_code);
                                                          printf("Flight %d (%s)\n    %d:%d%c.m. %s (%s)\n    %d:%d%c.m. %s (%s)\n", flight_number, travel_days, d_hour_stop, d_mins_stop, d_letter_stop, stop_city, stop_code, a_hour, a_mins, a_letter, arrival_city, arrival_input);
                                                        }
                                                    }

                                                }
                                            }
                                        }
                                    }
                                }

                              //if day entered is 4

                              if(input==4)
                                {
                                  while(fscanf(flights, "%d %d %d:%d%c %d:%d%c %s", flight_number, route_id, d_hour, d_mins, d_letter, a_hour, a_mins, a_letter, days) != EOF)
                                    {
                                      if(id==route_id)
                                        {
                                          while(z<7)
                                            {
                                              if(days[z]==4)
                                                {
                                                 // if(check_days() == TRUE)
                                                    {
                                                      //check_time();
                                                      if(layover >= 60 && layover <= 120)
                                                        {
                                                          printf("Flight %d (%s)\n    %d:%d%c.m. %s (%s)\n    %d:%d%c.m. %s (%s)\n", flight_number, travel_days, d_hour, d_mins, d_letter, departure_city, departure_input, a_hour_stop, a_mins_stop, a_letter_stop, stop_city, stop_code);
                                                          printf("Flight %d (%s)\n    %d:%d%c.m. %s (%s)\n    %d:%d%c.m. %s (%s)\n", flight_number, travel_days, d_hour_stop, d_mins_stop, d_letter_stop, stop_city, stop_code, a_hour, a_mins, a_letter, arrival_city, arrival_input);
                                                        }
                                                    }

                                                }
                                            }              
                                        }
                                    }
                                }

                              //if day entered is 5

                              if(input==5)
                                {
                                  while(fscanf(flights, "%d %d %d:%d%c %d:%d%c %s", flight_number, route_id, d_hour, d_mins, d_letter, a_hour, a_mins, a_letter, days) != EOF)
                                    {
                                      if(id==route_id)
                                        {
                                          while(z<7)
                                            {
                                              if(days[z]==5)
                                                {
                                                 // if(check_days() == TRUE)
                                                    {
                                                      //check_time();
                                                      if(layover >= 60 && layover <= 120)
                                                        {
                                                          printf("Flight %d (%s)\n    %d:%d%c.m. %s (%s)\n    %d:%d%c.m. %s (%s)\n", flight_number, travel_days, d_hour, d_mins, d_letter, departure_city, departure_input, a_hour_stop, a_mins_stop, a_letter_stop, stop_city, stop_code);
                                                          printf("Flight %d (%s)\n    %d:%d%c.m. %s (%s)\n    %d:%d%c.m. %s (%s)\n", flight_number, travel_days, d_hour_stop, d_mins_stop, d_letter_stop, stop_city, stop_code, a_hour, a_mins, a_letter, arrival_city, arrival_input);
                                                        }
                                                    }

                                                }
                                            }
                                        }
                                    }
                                }

                              //if day entered is 6

                              if(input==6)
                                {
                                  while(fscanf(flights, "%d %d %d:%d%c %d:%d%c %s", flight_number, route_id, d_hour, d_mins, a_hour, a_mins, a_letter, days) != EOF)
                                    {
                                      if(id==route_id)
                                        {
                                          while(z<7)
                                            {
                                              if(days[z]==6)
                                                {
                                                 // if(check_days() == TRUE)
                                                    {
                                                      //check_time();
                                                      if(layover >= 60 && layover <= 120)
                                                        {
                                                          printf("Flight %d (%s)\n    %d:%d%c.m. %s (%s)\n    %d:%d%c.m. %s (%s)\n", flight_number, travel_days, d_hour, d_mins, d_letter, departure_city, departure_input, a_hour_stop, a_mins_stop, a_letter_stop, stop_city, stop_code);
                                                          printf("Flight %d (%s)\n    %d:%d%c.m. %s (%s)\n    %d:%d%c.m. %s (%s)\n", flight_number, travel_days, d_hour_stop, d_mins_stop, d_letter_stop, stop_city, stop_code, a_hour, a_mins, a_letter, arrival_city, arrival_input);
                                                        }
                                                    }

                                                }                
                                            }
                                        }
                                    }
                                }

                              //if day entered is 7

                              if(input==7)
                                {
                                  while(fscanf(flights, "%d %d %d:%d%c %d:%d%c %s", flight_number, route_id, d_hour, d_mins, d_letter, a_hour, a_mins, a_letter, days) != EOF)
                                    {
                                      if(id==route_id)
                                        {
                                          while(z<7)
                                            {
                                              if(days[z]==7)
                                                {
                                                 // if(check_days() == TRUE)
                                                    {
                                                      //check_time();
                                                      if(layover >= 60 && layover <= 120)
                                                        {
                                                          printf("Flight %d (%s)\n    %d:%d%c.m. %s (%s)\n    %d:%d%c.m. %s (%s)\n", flight_number, travel_days, d_hour, d_mins, d_letter, departure_city, departure_input, a_hour_stop, a_mins_stop, a_letter_stop, stop_city, stop_code);
                                                          printf("Flight %d (%s)\n    %d:%d%c.m. %s (%s)\n    %d:%d%c.m. %s (%s)\n", flight_number, travel_days, d_hour_stop, d_mins_stop, d_letter_stop, stop_city, stop_code, a_hour, a_mins, a_letter, arrival_city, arrival_input);
                                                        }
                                                    }

                                                }
                                            }
                                        }
                                    }
                                }
                            }

                          /* if(input==0)
                      {
                        while(fscanf(flights, "%d, %d, (time) (time) (days)", flight_number, route_id, departure_time, arrival_time, (days)) != EOF)
                          {
                            if(id==route_id)
                              {
                                if(*days has daily*)
                                  {
                                    printf("Flight %d (%s)\n    %d:%d.m. %s, %s (%s)\n    %d:%d.m. %s, %s (%s)\n", flight_number, (days), d_hour, d_mins, departure_city, departure_state, departure_input, a_hour, a_mins, stop_city, stop_state, stop_code);
                                    printf("Flight %d (%s)\n    %d:%d.m. %s, %s (%s)\n    %d:%d.m. %s, %s (%s)\n", flight_number, (days), d_hour_stop, d_mins_stop, stop_city, stop_state, stop_code, a_hour_stop, a_mins_stop, arrival_city, arrival_state, arrival_input);
                                  }
                                else
                                  {
                                    printf("Route %s to %s has no 1-stop flights on the specified days\n", departure_input, arrival_input);
                                  }
                              }
                          }
                          }*/
                        }
                    }
                }
            }
        }
      else
        {
          printf("Either airport %s or %s doesn't exist", departure_input, arrival_input);
        }
    }
  return;
}
```